VM-Centric Snapshot Deduplication with Fault Isolation for Converged Cloud Architectures

### 2014-01-31 snapshot_system/paper_arch.tex
re

Our prototype system runs on a cluster of Linux machines with Xen-based VMs 
and the QFS~\cite{michael2013} distributd file system. 
%to %manage  the physical disk storage. 
All data needed for the backup service including snapshot data and metadata
%and snapshot data for backup purposes, 
resides in this distributed file system. 
One physical node hosts tens of VMs, each of which accesses its virtual machine disk image through the
virtual block device driver (called TapDisk\cite{Warfield2005} in Xen).

\subsection{Per Node Software Components} 
As depicted in Figure~\ref{fig:arch_vm}, 
there are four key service components running on each cluster
node  for supporting backup and deduplication: 
1) a virtual block device driver, 2) a snapshot deduplication agent,
3) a snapshot store client to store  and access snapshot data,
and 4)  a PDS client to support PDS metadata access.

@@

%\begin{figure}[htbp]
\begin{figure}[htbp]
    \centering
    %\includegraphics[width=2.75in]{images/socc_arch_cluster}
    \includegraphics[width=2.3in]{images/socc_arch_cluster}
    \caption{Per node architecture and data flow during snapshot backup}
    \label{fig:arch_vm}
\end{figure}

\subsection{A VM-centric Snapshot Store for Backup Data}
\label{sect:store}
We use the QFS distributed file system to hold snapshot backups.
Following the VM-centric idea for the purpose of fault isolation,
each VM has its own snapshot store, containing new data chunks which are considered
to be non-duplicates.
As shown in Figure~\ref{fig:as_arch}, we explain the data structure of the snapshot stores as follows.
%\begin{figure}[htbp]
\begin{figure}[htbp]
  \centering
%  \includegraphics[width=2.75in]{images/sstore_arch}
  \includegraphics[width=2.3in]{images/sstore_arch}
  \caption{VM snapshot store data structures}
  \label{fig:as_arch}
\end{figure}

%\begin{itemize}
%\item
There is an independent store containing all PDS chunks shared among different VMs as
a single file.
%All PDS chunks are stored in one PDS file. 
Each reference to a PDS data chunk in the PDS index is the offset within the PDS file.
%There is an option to use a data structure similar to the non-PDS store. We opt
%for the simpler format because
Additional compression is not applied because 
for the data sets we have tested, we only observed limited spatial locality 
among popular data chunks. On average the number of consecutive PDS index hits is lower than 7,
thus it is not very effective to group a large number of chunks as a compression and data fetch unit. 
For the same reason, we decide not to take the sampled index approach~\cite{Guo2011} 
for detecting duplicates from PDS as limited spatial locality is not sufficient to enable
effective prefetching for sampled indexing.

>>>>

Our prototype system runs on a cluster of Linux machines with Xen-based VMs 
and the QFS~\cite{michael2013} distributed file system. 
%to %manage  the physical disk storage. 
All data needed for the backup service including snapshot data and metadata
%and snapshot data for backup purposes, 
resides in this distributed file system. 
One physical node hosts tens of VMs, each of which accesses its virtual machine disk image through the
virtual block device driver (called TapDisk\cite{Warfield2005} in Xen).

\subsection{Per Node Software Components} 
%\begin{figure}[htbp]
\begin{figure}[htbp]
    \centering
    %\includegraphics[width=2.75in]{images/socc_arch_cluster}
    \includegraphics[width=2.2in]{images/socc_arch_cluster}
    \caption{Data flow during snapshot backup}
    \label{fig:arch_vm}
\end{figure}
As depicted in Figure~\ref{fig:arch_vm}, 
there are four key service components running on each cluster
node  for supporting backup and deduplication: 
1) a virtual block device driver, 2) a snapshot deduplication agent,
3) a snapshot store client to store  and access snapshot data,
and 4)  a PDS client to support PDS metadata access.

@@

\subsection{A VM-centric Snapshot Store for Backup Data}
\label{sect:store}
%\begin{figure}[htbp]
\begin{figure}[htbp]
  \centering
%  \includegraphics[width=2.75in]{images/sstore_arch}
  \includegraphics[width=2.2in]{images/sstore_arch}
  \caption{VM snapshot store data structures}
  \label{fig:as_arch}
\end{figure}

We use the QFS distributed file system to hold snapshot backups.
Following the VM-centric idea for the purpose of fault isolation,
each VM has its own snapshot store, containing new data chunks which are considered
to be non-duplicates.
As shown in Figure~\ref{fig:as_arch}, we explain the data structure of the snapshot stores as follows.
%\begin{itemize}
%\item
There is an independent store containing all PDS chunks shared among different VMs as
a single file.
%All PDS chunks are stored in one PDS file. 
Each reference to a PDS data chunk in the PDS index is the offset within the PDS file.
%There is an option to use a data structure similar to the non-PDS store. We opt
%for the simpler format because
Additional compression is not applied because 
for the data sets we have tested, we only observed limited spatial locality 
among popular data chunks. On average the number of consecutive PDS index hits is lower than 7,
thus it is not very effective to group a large number of chunks as a compression and data fetch unit. 
For the same reason, we decide not to take the sampled index approach~\cite{Guo2011} 
for detecting duplicates from PDS as limited spatial locality is not sufficient to enable
effective prefetching for sampled indexing.

### 2014-01-31 snapshot_system/paper_eval.tex

Table~\ref{tab:leakage} shows the average accumulated storage leakage in terms of percentage of
storage space per VM caused  by approximate deletions.
In this experiment, we select 1050 snapshots from 105 VMs and let all the VMs accumulate 10 snapshots, 
then start to delete those snapshots one by one.
As we know the actual storage needs after each snapshot creation,
the storage leakage can be detected by
comparing the size of remaining data in use after deletion to the correct number.
The top dashed line is the predicted leakage using Formula~\ref{eq:leakrepair} from Section~\ref{sect:delete}
given $\Delta u/u=0.025$,
while the solid line represents the actual average leakage measured during the experiment for all the VMs. 
The Bloom filter setting is based on $\Delta u/u=0.025$.
After 9 snapshot deletions, the actual leakage ratio reaches 0.0015 and this means that
there is only 1.5MB space leaked for every 1GB of stored data.
The actual leakage can reach  4.1\% after  245 deletions and such a repair is needed every 12 days
to the  leakage under 5\%. 
This experiment shows that the leakage of our approximate snapshot deletion is very small, 
below the estimated number.

The space cost for deletion is insignificant.
Leakage repair needs less than 85MB of memory
considering each reference consumes 8 bytes plus  1 mark bit and  
each VM snapshot has 40GB backup data with about 10 million chunks. 
This VM-specific leakage  repair takes  less than half an hour for each VM, 
when disk bandwidth is controlled to 50MB/s. 
%When using the traditional mark-and-sweep~\cite{Guo2011,Fabiano2013} in VO for 2500 VMs, 
%a total of 212GB memory
%space on 100 node is accessed for processing each VM and each machine requires 2.12GB.
%To control the garbage growth, such a process needs to be conducted much more frequently than our
%VM-centric leakage repair.

%Assuming snapshot deletion happens daily,
%such a  mark-and-sweep deletion needs to be conducted every two days to control garbage space under 5\%
%since there is 2.5\% data change for each snapshot.

% We also measure the storage leakage under $\Delta u/u=0.05$ and $0.10$ settings for summary vectors.
% As this value increases, larger summary vectors are generated thus results in a decrease on the
% false-positive ratio, therefore less amount of unused chunks are misjudged.

>>>>

Table~\ref{tab:leakage} shows the average accumulated storage leakage in terms of percentage of
storage space per VM caused  by approximate deletions.
In this experiment, we select 105 VMs and let all the VMs accumulate 10 snapshots, 
then start to delete those snapshots one by one in reverse order.
As we know the actual storage needs after each snapshot creation,
the storage leakage can be detected by
comparing the size of remaining data in use after deletion to the correct number.
The top dashed line is the predicted leakage using Formula~\ref{eq:leakrepair} from Section~\ref{sect:delete}
given $\Delta u/u=0.025$,
while the solid line represents the actual average leakage measured during the experiment for all the VMs. 
The Bloom filter setting is based on $\Delta u/u=0.025$.
After 9 snapshot deletions, the actual leakage ratio reaches 0.0015 and this means that
there is only 1.5MB space leaked for every 1GB of stored data.
The actual leakage can reach  4.1\% after  245 deletions and such a repair is needed every 12 days
to the  leakage under 5\%. 
This experiment shows that the leakage of our approximate snapshot deletion is very small, 
below the estimated number.
